---
title: "housing_formulas.rmd"
output: html_document
date: "2023-04-30"
---

```{r}
install.packages("tidyverse")
install.packages("gapminder")

```

```{r}
install.packages("writexl")
```
```{r}
install.packages(ggplot2)
```

```{r}
library(tidyverse)
library(gapminder)
library(readxl)
library(writexl)
library(ggplot2)
```

```{r}
miami_average <- read_excel("Miami_Dade_Average_2018-2022.xlsx")
miami_average
```

```{r}
broward_average <- read_excel("Broward_Average_2018_2022.xlsx")
broward_average
```

```{r}
palm_average <- read_excel("Palm_Beach_Average_2018-2022.xlsx")
palm_average
```

#Miami as a county and usuable 
```{r}
miami_pivot <- miami_average %>% pivot_longer(cols=-1)
colnames(miami_pivot)[1] <- "Date"
colnames(miami_pivot)[2] <- "Year_of_Sale"
colnames(miami_pivot)[3] <- "Average_Price"
miami_pivot$County <- rep("Miami_Dade", nrow(miami_pivot))
```

#Broward 
```{r}
bro_pivot <- broward_average %>% pivot_longer(cols=-1)
colnames(bro_pivot)[1] <- "Date"
colnames(bro_pivot)[2] <- "Year_of_Sale"
colnames(bro_pivot)[3] <- "Average_Price"
bro_pivot$County <- rep("Broward", nrow(bro_pivot))
```

#PalmBeach 
```{r}
palm_pivot <- palm_average %>% pivot_longer(cols=-1)
colnames(palm_pivot)[1] <- "Date"
colnames(palm_pivot)[2] <- "Year_of_Sale"
colnames(palm_pivot)[3] <- "Average_Price"
palm_pivot$County <- rep("Palm_Beach", nrow(palm_pivot))
```

#Combining all new datasets into one 
```{r}
combined_counties <- rbind(bro_pivot, palm_pivot, miami_pivot)
combined_counties$Date <- factor(combined_counties$Date, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
```

```{r}
combined_counties$Year_of_Sale <- gsub("Average_Sale_Price_2018", "2018", combined_counties$Year_of_Sale)
```

```{r}
combined_counties$Year_of_Sale <- gsub("Average_Sale_Price_2019", "2019", combined_counties$Year_of_Sale)

combined_counties$Year_of_Sale <- gsub("Average_Sale_Price_2020", "2020", combined_counties$Year_of_Sale)

combined_counties$Year_of_Sale <- gsub("Average_Sale_Price_2021", "2021", combined_counties$Year_of_Sale)

combined_counties$Year_of_Sale <- gsub("Average_Sale_Price_2022", "2022", combined_counties$Year_of_Sale)

combined_counties$Year_of_Sale <- gsub("Average_Sale_Price_2023", "203", combined_counties$Year_of_Sale)
```


#exporting the excel as one 
```{r}
write_xlsx(combined_counties, "Counties_Average.xlsx")


a <- 1:12
a
```

#Now creating the graph that will show the different prices by year 
```{r}
p <- ggplot(data = combined_counties, 
            mapping = aes(x = Date, 
                          y = Average_Price)) + 
  geom_point(mapping = aes(color = Year_of_Sale)) +
  labs(title = "Average Price Per Month",
       x = "Month",
       y = "Price")
p
```




#Next table to be done 
```{r}
broward <- read_excel("LuxuryHousesBroward.xlsx")
broward
```

#Give dbl to the columns ill use 
```{r}
broward$City <- factor(broward$City)
broward$Sold_Price <- as.double(broward$Sold_Price)
broward$Sold_Date <- factor(broward$Sold_Date)
broward$Bedrooms <- as.double(broward$Bedrooms)
```

#Creating the data set with the infor I will actually use 
```{r}
broward_county <- broward[,c(1,3,11,12,23)]
colnames(broward_county)[1] <- "Listing_Number"
colnames(broward_county)[2] <- "City"
colnames(broward_county)[3] <- "Price"
colnames(broward_county)[4] <- "Date"
colnames(broward_county)[5] <- "Bedrooms"
```

```{r}
broward_county$Year <- year(as.Date(broward_county$Date))
broward_county <- select(broward_county, -Date)
```


```{r}
write_xlsx(broward_county, "broward_file.xlsx")
```


```{r}
library(ggplot2)
```


#By Bedrooms
```{r}
facet <- ggplot(data = broward_county, 
            mapping = aes(x = Year, 
                          y = Price)) + 
  geom_point(mapping = aes(color = factor(Bedrooms))) + 
 scale_y_log10(labels = scales::dollar) +
  theme(axis.text.x=element_text(angle=90)) +
  facet_wrap(~City)
facet
```












#Last graph 
```{r}
water <- read_excel("WaterFront_Broward.xlsx")
water
```



#Deleting the yyyy-mm-dd column and making it a year only column 
```{r}
water$Year <- year(as.Date(water$Sold_Date))
water <- select(water, -Sold_Date)
```

```{r}
water <- water[!(water$Year %in% c(2018, 2019)),]
```



```{r}
plot <- ggplot(data = water,
            mapping = aes(x = Year,
                          y = Sold_Price,
                          fill = Water_Front)) + 
            geom_bar(stat = "identity", position = "dodge") + 
            theme(legend.position = "top") +
            scale_y_log10(labels = scales::dollar)
plot
```



# I am going to find the average price of houses with front or no by year by county 
```{r}
yes <- water[water$Water_Front == "Y", ]
no <- water[water$Water_Front == "N", ]
```


#Separating yes and no by city 
```{r}
yes_fl <- yes[yes$City == "Fort Lauderdale", ]
yes_hll <- yes[yes$City == "Hollywood", ]
yes_ppb <- yes[yes$City == "Pompano Beach", ]
yes_wm <- yes[yes$City == "Wilton Manors", ]
yes_lbts <- yes[yes$City == "Lauderdale By The Sea", ]
```


```{r}
no_fl <- no[no$City == "Fort Lauderdale", ]
no_hll <- no[no$City == "Hollywood", ]
no_ppb <- no[no$City == "Pompano Beach", ]
no_wm <- no[no$City == "Wilton Manors", ]
no_lbts <- no[no$City == "Lauderdale By The Sea", ]
```

```{r}
library(magrittr)
```

```{r}
library(dplyr)

```

#Find average prices by years by waterfront and adding the yes column and city column 
```{r}
yes_fl_price <- yes_fl %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  yes_fl_price$Water_Front <- rep("Y", nrow(yes_fl_price))
  yes_fl_price$City <- rep("Fort Lauderdale", nrow(yes_fl_price))

yes_hll_price <- yes_hll %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  yes_hll_price$Water_Front <- rep("Y", nrow(yes_hll_price))
  yes_hll_price$City <- rep("Hollywood", nrow(yes_hll_price))

yes_ppb_price <- yes_ppb %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  yes_ppb_price$Water_Front <- rep("Y", nrow(yes_ppb_price))
  yes_ppb_price$City <- rep("Pompano Beach", nrow(yes_ppb_price))

yes_wm_price <- yes_wm %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  yes_wm_price$Water_Front <- rep("Y", nrow(yes_wm_price))
  yes_wm_price$City <- rep("Wilton Manors", nrow(yes_wm_price))

yes_lbts_price <- yes_lbts %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  yes_lbts_price$Water_Front <- rep("Y", nrow(yes_lbts_price))
  yes_lbts_price$City <- rep("Lauderdale By The Sea", nrow(yes_lbts_price))

```


```{r}
no_fl_price <- no_fl %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  no_fl_price$Water_Front <- rep("N", nrow(no_fl_price))
  no_fl_price$City <- rep("Fort Lauderdale", nrow(no_fl_price))

no_hll_price <- no_hll %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  no_hll_price$Water_Front <- rep("N", nrow(no_hll_price))
  no_hll_price$City <- rep("Hollywood", nrow(no_hll_price))

no_ppb_price <- no_ppb %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  no_ppb_price$Water_Front <- rep("N", nrow(no_ppb_price))
  no_ppb_price$City <- rep("Pompano Beach", nrow(no_ppb_price))

no_wm_price <- no_wm %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  no_wm_price$Water_Front <- rep("N", nrow(no_wm_price))
  no_wm_price$City <- rep("Wilton Manors", nrow(no_wm_price))

no_lbts_price <- no_lbts %>% 
  group_by(Year) %>% 
  summarize(avg_price = mean(Sold_Price))
  no_lbts_price$Water_Front <- rep("N", nrow(no_lbts_price))
  no_lbts_price$City <- rep("Lauderdale By The Sea", nrow(no_lbts_price))
```

#Combining all the average prices in one file 
```{r}
broward_cities <- rbind(no_lbts_price, no_wm_price, no_ppb_price, no_hll_price, no_fl_price, yes_lbts_price, yes_wm_price, yes_ppb_price, yes_hll_price, yes_fl_price)

```

```{r}
write_xlsx(broward_cities, "water.xlsx")
```


```{r}
write_xlsx(water, "3tab.xlsx")
```

